---
title: "Using azmetr with the units package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using azmetr with the units package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(azmetr)
library(units)
library(dplyr)
```

## Adding units to data

You can add the correct units to data returned by `az_daily()`, `az_hourly()`, or `az_heat()` by passing the resulting tibble to `az_add_units()`.

```{r}
hourly <- 
  az_hourly() |> 
  az_add_units() 

hourly |> 
  select(-starts_with("meta_"), -starts_with("date_")) |> 
  head()
```

This requires that you have the `units` package installed and will prompt you to do so if you don't have it installed.  It may also be helpful to explicitly load the package with `library(units)` so that the resulting tibble displays the units correctly. 

## Using units columns

`az_add_units()` converts numeric vectors to those of class "units".  These units columns behave differently than ordinary numeric vectors and have a few useful properties. First, you can do unit conversion using `set_units()` from the `units` package.

```{r}
hourly |> 
  transmute(wind_spd_kph = set_units(wind_spd_mps, "km/h"),
            sol_rad_total = set_units(sol_rad_total, "W h m-2"),
            temp_airK = set_units(temp_airF, "Kelvins"))
```

Second, it won't allow you to do math where the units aren't compatible.

```{r}
#| error: true
hourly |> 
  transmute(wind_rain = wind_spd_mps + precip_total)
```

That also means that you generally cannot add or subtract unitless constants.

```{r}
## This will error:
# hourly$wind_spd_mps[1] + 10

## Must use:
hourly$wind_spd_mps[1] + set_units(10, "m/s")
```


## Plotting with units

The `units` package works with `ggplot2` to automatically include units in axis labels.

```{r}
library(ggplot2)
ggplot(hourly, aes(x = wind_spd_mps, y = sol_rad_total)) +
  geom_point() +
  labs(x = "wind speed",
       y = "total solar radiation")
```

