---
title: "Creating weather maps with {azmetr} and {ggplot2}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating weather maps with {azmetr} and {ggplot2}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Load packages

```{r setup}
library(azmetr)
library(ggplot2)
library(dplyr)
library(maps)
```

# Get map data

We can use the `maps` package to get data for drawing the state of Arizona

```{r}
az <- map_data("state", "arizona")
p <- 
  ggplot() +
  geom_polygon(data = az, aes(x = long, y = lat, group = group), color = "black", fill =  "lightblue") +
  coord_map()
p
```

# Get weather data

Use `az_daily()` without any arguments to retrieve the most recent daily average for each AZMET station.

```{r}
weather <- az_daily()
head(weather)
```

To attach location info to this data we can join it to the `station_info` dataset

```{r}
df <- left_join(weather, station_info)
colnames(df)
```

Now each row has a `latitude` and `longitude` column that can be used for adding points or labels.

```{r}
p + 
  geom_point(data = df, aes(x = longitude, y = latitude, color = temp_air_meanC))

#temp as text
p +
  geom_text(data = df, aes(x = longitude, y = latitude, label = temp_air_meanC))

#station ID as text and temp as color
p +
  geom_text(data = df, aes(x = longitude, y = latitude, label = meta_station_id, color = temp_air_meanC))
```
Finally, we can do some styling to improve the appearance of the map and add a title.

```{r}
p + 
  geom_point(data = df, aes(x = longitude, y = latitude, color = temp_air_meanC)) +
  scale_color_viridis_c(option = "C") +
  theme_void() +
  labs(
    title = df$datetime |> unique(),
    color = "Mean air temp (ºC)"
  )
```

